---
title: "Introduction to R"
subtitle: "For Statistical Analysis"
author: "John Palmer"
date: "November 15, 2019"
output: 
  revealjs::revealjs_presentation:
    theme: beige
    highlight: pygments
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introducing R

## What is R?

- R is a programming language and software environment for statistical computing and graphics
- It comes from the S language developed by John Chambers at Bell Laboratories in the 1970s
- R was created by Ross Ihaka and Robert Gentleman at the University of Auckland, New Zealand, starting in the 1990s. It is now developed by a larger community of developers (the R Development Core Team) supported by non-profit foundation (the R Foundation)
- Official home page is https://r-project.org/

## R is Free Software

- Being "free software" means users have the freedom to run, copy, distribute, study, change and improve R
- In other words, the "free" part is about liberty not price
- But R is also free in terms of price

## R Packages

- One of the powerful thing about R is the enormous ecosystem of packages that people have written for it
- An R package is a bundle of code, often along with data, documentation and tests, which runs in R and can be shared
- There are thousands of packages for R, written by people all over the world in a wide variety of disciplines, and mostly distributed under free and open source licenses
- You can easily install R packages from CRAN as well from GitHub and other code repositories


## Installing R

- You should already have it installed, but if not, go to https://cran.r-project.org/ and install the latest release that is compatible with your system

## Upgrading R

- You do not need to upgrade every time a new version is released, but you will want to upgrade every so often in order to take advantage of new features or to ensure support for new packages
- If you already have numerous packages, you may want to copy them into the new version to avoid having to download them again
- It is sometimes useful to have multiple versions of R installed side-by-side and this is possible

# Working with R

## Terminal

- The Read-Eval-Print Loop is the most basic way of interacting with R directly
- You can also run R files (non-interactively) from a terminal

## Text Editor

- You can use any text editor to create, store, and edit R scripts (like do-files)
- It is nice to have features like code highlighting and a built-in REPL.
- Choices include Emacs, vim, and sublime

## IDE

- An Integrated Development Environment combines a text editor with a set of  other useful tools for debugging, automation, and other tasks
- The open source version of R Studio is the obvious choice here

# R Basics

## Basic Data Types

- character: `r "a"`
- numeric: `r 2` `r 2.4`
- integer: `r 2L` 
- logical: `r TRUE`
- complex: `r 1+4i`

## Assignment

```{r, echo=TRUE, eval=TRUE}

x <- 1
y <- 2
z <- x + y
print(z)
```

## More Data Types

- https://raw.githubusercontent.com/rstudio/cheatsheets/master/base-r.pdf

## Key Concepts

- Based and Packages
- Object Oriented Programming
- Naming things and understanding environments

# Statistics with R


